A radiocarbon data based evaluation of Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe

### 2019-01-26 article/03_Results.Rmd
language in results

In the course of the Bronze Age in Europe there was a trend away from body burial towards cremation. Around 2200calBC cremation burials were almost unknown in Poland, southern Germany, northeastern France and northern Germany. In the Northwest, in Great Britain and the Benelux area, and in the Southeast, in Austria and the Czech Republic, however, there were early cremation traditions. These contexts could be discussed as areas of origin of the later omnipresent phenomenon. While cremation burials lost importance in Austria and the Czech Republic as well as in the Benelux region during the Early Bronze Age and only regained their importance in the Middle Bronze Age, their share increased steadily in England. Inhumation and cremation were balanced here for a long time. The situation was similar in southern Scandinavia, where the proportion of cremation burials was much lower until the Late Bronze Age. In Poland and northern Germany, cremation became relatively suddenly and completely established with the beginning of the Late Bronze Age. In northeastern France and northern Germany this change took place already in the Middle Bronze Age. In both contexts, inhumation continued to play an important role.

With regard to the question of grave mound construction, the picture is considerably more heterogeneous. In Austria and the Czech Republic, flat graves were the dominant form of burial until the Late Bronze Age, but mounds became more relevant from the Middle Bronze Age onwards. In Poland, flat and hill graves remained in balance until the Middle Bronze Age, when flat graves prevailed. The amount of data from southern Germany and northeastern France is extraordinarily small: If one believes the sample, then a sudden, radical change from burial in flat graves to hill graves took place in southern Germany at the beginning of the Middle Bronze Age. In northeastern France there were no flat graves in the Bronze Age. Northern Germany underwent a development from the dominance of flat graves in the Early Bronze Age, a short phase of increasing mound construction in the Middle Bronze Age followed by renewed dominance of the flat grave custom in the Late Bronze Age. A similar development can be seen in southern Scandinavia: Flat graves clearly predominated, but were increasingly -- almost completely -- replaced by hills, until burial in flat graves became more frequent again at the end of the Bronze Age. Burial mounds were dominant throughout the Benelux region, but in the Early and Late Bronze Ages flat graves also occurred to a significant extent. In England, flat graves were a short phenomenon in the Early Bronze Age, which did not recur later. \newline

@@

The results of the simulation as presented in Figure \ref{simulation} confirm the expectation: In case of equal interaction intensity $I$ of separate populations, the correlation of cultural and spatial distance is weak (Figure \ref{simulation} A vs. C vs. E). It appears to be normally distributed around zero for many simulation runs, although individual runs can still yield remarkably high positive and negative correlation for certain time slots. The fluctuation of the correlation within one run can be strong and seems to increase for bigger values of $m_i$. If the interaction matrix $I$ is predefined to represent the spatial distance matrix (Figure \ref{simulation} BDF), then the mean correlation per time slot increases. The correlation increases as well with the degree of interaction $m_i$ (Figure \ref{simulation} D vs. F). Counterintuitively the fluctuation of the correlation is also here very high (Figure \ref{simulation} E vs. F). The correlation development observed for the variables *burial type* and *burial construction* fits to simulation runs in all presets, but becomes less likely if we assume $I$ to represent spatial distance and a high degree of interaction $m_i$ (Figure \ref{simulation} A vs. F). That's mostly because of the relatively strong negative correlation of cultural and spatial distance in the Middle and Late Bronze Age (1400 to 1000calBC).

>>>>

In the course of the Bronze Age in Europe there was a trend away from body burial towards cremation. Around 2200calBC cremation burials were almost unknown in Poland, Southern Germany, Northeastern France and Northern Germany. In the Northwest, in Great Britain and the Benelux area, and in the Southeast, in Austria and the Czech Republic, however, there were early cremation traditions. These contexts could be discussed as areas of origin of the later omnipresent phenomenon. While cremation burials lost importance in Austria and the Czech Republic as well as in the Benelux region during the Early Bronze Age and only regained relevance there in the Middle Bronze Age, their share increased steadily in England. Inhumation and cremation were balanced here for a long time. The situation was similar in Southern Scandinavia, where the proportion of cremation burials was much lower until the Late Bronze Age. In Poland and northern Germany, cremation became relatively suddenly and completely established with the beginning of the Late Bronze Age. In Northeastern France and Northern Germany this change took place already in the Middle Bronze Age. In both contexts inhumation continued to play an important role.

With regard to the question of grave mound construction the picture is considerably more heterogeneous. In Austria and the Czech Republic flat graves were the dominant form of burial until the Late Bronze Age, but mounds became more relevant from the Middle Bronze Age onwards. In Poland mounds and flat graves remained in balance until the Middle Bronze Age, when the latter finally prevailed. The amount of data from Southern Germany and Northeastern France is extraordinarily small: If one believes the sample, then a sudden, radical change from burial in flat graves to hill graves took place in Southern Germany at the beginning of the Middle Bronze Age. In Northeastern France there were no flat graves in the Bronze Age. Northern Germany underwent a development from the dominance of flat graves in the Early Bronze Age, a short phase of increasing mound construction in the Middle Bronze Age followed by renewed dominance of the flat grave custom in the Late Bronze Age. A similar development can be observed in Southern Scandinavia: Flat graves clearly predominated, but were increasingly -- almost completely -- replaced by mounds, until burial in flat graves became more frequent again at the end of the Bronze Age. Burial mounds were dominant throughout the Benelux region, but in the Early and Late Bronze Ages flat graves also occurred to a significant extent. In England flat graves were a short phenomenon in the Early Bronze Age, which did not recur later. \newline

@@

The results of the simulation as presented in Figure \ref{simulation} confirm the expectation: In case of equal interaction intensity $I$ of separate populations, the correlation of cultural and spatial distance is weak (Figure \ref{simulation} A vs. C vs. E). It appears to be normally distributed around zero for many simulation runs, although individual runs can still yield remarkably high positive and negative correlation for certain time slots. The fluctuation of the correlation within one run can be strong and seems to increase for bigger values of $m_i$. If the interaction matrix $I$ is predefined to represent the spatial distance matrix (Figure \ref{simulation} BDF), then the mean correlation per time slot increases. The correlation increases as well with the degree of interaction $m_i$ (Figure \ref{simulation} D vs. F). Counter-intuitively the fluctuation of the correlation is also here very high (Figure \ref{simulation} E vs. F). The correlation development observed for the variables *burial type* and *burial construction* fits to simulation runs in all presets, but becomes less likely if we assume $I$ to represent spatial distance and a high degree of interaction $m_i$ (Figure \ref{simulation} A vs. F). That's mostly because of the relatively strong negative correlation of cultural and spatial distance in the Middle and Late Bronze Age (1400 to 1000calBC).
